Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-06-26 paper/paper.Rmd
optimize discussion

While in the example the `"spcv_coords"` method was used, this should not imply that this method is the best or only method suitable for this example task.
Even though this method is frequently used, it was mainly chosen because of it's clear visual grouping differences when comparing spatial to random partitioning.

In fact, most often multiple spatial partitioning methods can be used for a dataset.
It is recommended (required) that users familiarize themselves with multiple methods and decide which method to choose based on the specific characteristics of the dataset at hand.
There are various variables influencing the partitioning method choice for an environmental dataset:

@@

Based on these criteria users must choose a matching partitioning method that is either more restrictive (by discarding observations for fold creation) or more liberal (by using all observations and eventually ignoring natural grouping patterns).
For almost all methods implemented in \pkg{mlr3spatiotempcv}, there is a scientific publication describing the strengths and weaknesses of the respective approach (either linked in the help file of the package or its respective upstream packages).

In the example above, a cross-validation without hyperparameter tuning was shown.
In practice most often a nested CV is desired.
In this case it is recommended to use the same spatial partitioning method for the inner loop (= tuning level).
See @schratz2019 for more details and chapter 11 of [Geocomputation with R](https://geocompr.robinlovelace.net/spatial-cv.html) [@lovelace2019].

## Comparing spatiotemporal resampling methods

<!-- high autocor on the border of test sets -->
<!-- sptcv LLTO: discards many observations -->
All methods which make use of all variables in the dataset suffer from the problem that on the border between train and test sets, both in the spatial and temporal domain, autocorrelation is quite high.
Only methods that omit some observations (`"spcv_buffer"`, `"sptcv_cstf"` with both `space_var` and `time_var` set) are able to reduce this problem.
However, these then suffer from the fact of potentially not being able to properly represent the dataset because only a subset of observations is used.

<!-- sptcv LLO & LTO: requires additional variable to specify groupings -->
Methods `"spcv_cstf"` (with only one of `space_var` or `time_var` set) require a variable in the dataset which should be used for grouping.
The specification of such is not always easy because the selection of such a variable also implies that the selected groups inherit substantial autocorrelation and differ substantially from other groups.
Also, if the variable contains too many groups, the difference within train/test splits may become undesirably high and tend towards a LOO CV.

<!-- only some methods are suited for repeated execution -->
<!-- spcv_coords: limited variance due to limited k-means possibilities -->
For repeated executions of method `spcv_coords`, the variance between repetitions could be quite small due to the limited possibilities of different `k-means` clusters on the same dataset.
This applies even more to `"spcv_block"` with options `selection = "systematic"` and `selection = "checkerboard"` as the fold assignments are the same in each repetition.
Only methods with a random component (e.g. `"spcv_coords"` or `"spcv_block"` with option `selction = "random"`) may be suited for using multiple repetitions.

<!-- spcv buffer: discards observations, LOO very expensive -->
Method `"spcv_buffer"` makes use of the removal of observations to reduce the influence of spatial autocorrelation between train and test set.
This might be pretty effective in reducing autocorrelation but also requires to discard many variables and comes with the downsides of LOO-CV: many model fits which can be computationally very demanding and a high variance due to the high similarity of training datasets [@blockCV].

<!-- spcv coords: mention generic name -->
In @brenning2012, which is commonly used as the main reference for this k-means based spatial clustering approach, the method was not assigned a unique name.
This brings potential confusion for users and the community as the term "spatial CV" is used a generic one for the idea of accounting for spatial autocorrelation for spatial datasets.
Hence, in \pkg{mlr3spatiotempcv} this method was named `"spcv_coords"`, making use of the fact that the spatial coordinates are used for a cluster-based partition creation.

<!-- env block: hard to decide which variable to use for clustering -->
<!-- env block: more than one variable may lead to unforeseeable results -->
Method `"spcv_env"` is complicated in that it requires the user to limit the clustering to (a) specific variable(s).
This implies that only this/these variable(s) suffer from autocorrelation and others can be ignored.
However, the influence of a single variable is hard to estimate and the same applies for estimating the influence of the non-accounted variables which might add substantial bias to the CV.

@@

```{r echo=FALSE}
sessionInfo()
```

>>>>

The question which resampling method should be chosen for a dataset at hand comes up regularly in practice.
Even though there is no definitive and easy answer, we would like to give some guidance in this section to help finding the most appropriate method.

Although the case study example in section \@ref(sec:case-study) used the `"spcv_coords"` method, this should not imply that this method is the best or only method suitable for this example task.
Even though this method is frequently used, it was mainly chosen because of it's clear visual grouping differences when comparing spatial to random partitioning.

In fact, most often multiple spatial partitioning methods can be used for a dataset.
It is recommended (required) that one familiarizes themselves with multiple methods and decide which method to choose based on the specific characteristics of the dataset at hand.
There are various variables influencing the partitioning method choice for an environmental dataset:

@@

Based on these criteria users must choose a matching partitioning method that is either more restrictive (by discarding observations for fold creation) or more liberal (by using all observations and eventually ignoring natural grouping patterns).
For almost all methods implemented in \pkg{mlr3spatiotempcv}, there is a scientific publication describing the strengths and weaknesses of the respective approach (either linked in the help file of the package or its respective upstream packages), hence we are refraining from repeating them again in this study.

In the example above, a cross-validation without hyperparameter tuning was shown.
In ML practice most often a nested CV should be desired to optimize model hyperparameter within the CV.
In such scenarios it is recommended to use the same spatial partitioning method for the inner loop (= tuning level) that was chosen for the outer one.
See @schratz2019 for more details as well as chapter 11 of [Geocomputation with R](https://geocompr.robinlovelace.net/spatial-cv.html) [@lovelace2019].

## Comparing and discussing spatiotemporal resampling methods

<!-- high autocor on the border of test sets -->
<!-- sptcv LLTO: discards many observations -->
All methods which make use of all variables in the dataset suffer from the problem that on the border between train and test sets, both in the spatial and temporal domain, autocorrelation is relatively high (unless the dataset inherits natural gaps between train and test sets).
Only methods that omit some observations (`"spcv_buffer"`, `"spcv_tiles"`, `"sptcv_cstf"` with both `space_var` and `time_var` set) are able to reduce the impact of this issue.
However, these then suffer from the fact of potentially not being able to properly represent the dataset because only a subset of observations is used.

<!-- sptcv LLO & LTO: requires additional variable to specify groupings -->
The `"spcv_cstf"` methods (with only one of `space_var` or `time_var` set) require a variable in the dataset which should be used for grouping.
The specification of such is not always easy because the grouping should ensure that the selected groups inherit substantial autocorrelation within themselves and simultaneously differ substantially from other groups.
Also, if the variable selected contains too many groups, the difference within train/test splits may become undesirably high and tend towards a LOO CV [@meyer2018].

<!-- only some methods are suited for repeated execution -->
<!-- spcv_coords: limited variance due to limited k-means possibilities -->
For repeated executions of method `spcv_coords`, the variance between repetitions can become quite small due to the limited possibilities of differing `k-means`-based clusters on the same dataset [@brenning2012].
This applies even more to `"spcv_block"` with options `selection = "systematic"` and `selection = "checkerboard"` as the fold assignments are the same in each repetition.
In general, only methods making use of a random component (e.g. `"spcv_coords"` or `"spcv_block"` with option `selction = "random"`) may be suited for using multiple repetitions.

<!-- spcv buffer: discards observations, LOO very expensive -->
Method `"spcv_buffer"` makes use of the removal of observations to reduce the influence of spatial autocorrelation between train and test set.
This can be a pretty effective way to reduce autocorrelation but also requires to discard many observations and comes with the downsides of a LOO-CV: many model fits which can be computationally very demanding and the results inheriting a high variance due to the high differences in the test sets (also commonly known as the "bias-variance tradeoff" [@wahba1998; @blockCV].

<!-- spcv coords: mention generic name -->
In @brenning2012, which is the reference commonly used for the k-means based spatial partitioning approach, the method was not introduced with a unique name.
Over time, studies which have used this method mainly referenced it using the generic "spatial CV" terminology.
However, using this notation potentially introduces confusion as the term "spatial CV" is a generic one referring to the idea of accounting for spatial autocorrelation in spatial datasets and does not necessarily refer to a specific partitioning method in the first place.
Hence, in package \pkg{mlr3spatiotempcv} this method was named `"spcv_coords"`, making use of the fact that the spatial coordinates are used for a cluster-based partition creation.

<!-- env block: hard to decide which variable to use for clustering -->
<!-- env block: more than one variable may lead to unforeseeable results -->
Method `"spcv_env"` is complicated to use in that it requires the user to limit the clustering to one or more variables [@blockCV].
This implies that only this/these variable(s) suffer from autocorrelation and others are ignored.
However, it is not possible to estimate how much a single variable accounts for the autocorrelation in the data and the same applies for estimating the influence of the non-accounted variables which might add substantial bias to the CV.

@@

```{r echo=FALSE, cache=FALSE}
sessionInfo()
```
